Identifying the L1 of non-native writers: the CMU-Haifa system\\[1em] {\large Yulia Tsvetkov$^\ast$\ Naama Twitto$^\dagger$\ Nathan Schneider$^\ast$\ Noam Ordan$^\dagger$\\ Manaal Faruqui$^\ast$\ Victor Chahuneau$^\ast$\ Shuly Wintner$^\dagger$ \ Chris Dyer$^\ast$}\\[-3em]

### 2013-04-10 paper/paper.tex
use official NAACL bib style and other tweaks

\usepackage[round]{natbib}
\setlength\titlebox{6.5cm}    % Expanding the titlebox

@@

\newcommand{\feat}[1]{\textsmaller[.5]{\textsf{#1}}} % code for a feature group
\newcommand{\featstr}[1]{\textsmaller[.5]{\verb|#1|}} % full feature name
\newcommand{\textnl}{\textsl}

@@

\section{Related work}
\label{sec:literature}
The task of L1 identification was introduced by
\citet{koppel2005automatically,koppel2005determining}, who work on the
International Corpus of Learner English \citep{icle}, which includes
texts written by students from 5 countries, Russia, the Czech Republic, Bulgaria,
France, and Spain. The texts range from 500~to 850~words in length. Their
classification method is a linear SVM, and features include 400~standard function words, 200~letter $n$-grams, 185~error types and 250
rare part-of-speech (POS) bigrams. Ten-fold cross-validation results
on this dataset are 80\% accuracy.

The same experimental setup is assumed by
\citet{tsur-rappoport:2007:Cognitive-2007}, who are mostly interested
in testing the hypothesis that an author's choice of words in a second
language is influenced by the \emph{phonology} of his or her L1. They confirm
this hypothesis by carefully analyzing the features used by
\citeauthor{koppel2005automatically}, controlling for potential
biases.

\citet{Wong-Dras:2009:ALTA2009,wong-dras:2011:EMNLP} are also
motivated by a linguistic hypothesis, namely that \emph{syntactic} errors in
a text are influenced by the author's
L1. \citet{Wong-Dras:2009:ALTA2009} analyze three error types
statistically, and then add them as features in the same experimental
setup as above (using LIBSVM with a radial kernel for
classification). The error types are subject-verb disagreement,
noun-number disagreement and misuse of determiners. Addition of these
features does not improve on the results of
\citeauthor{koppel2005automatically}. \citet{wong-dras:2011:EMNLP}
further extend this work by adding as features horizontal slices of
parse trees, thereby capturing more syntactic structure. This improves
the results significantly, yielding 78\% accuracy compared with less
than 65\% using only lexical features.

@@

\section{Model}
\label{sec:model}
For our classification model we used the \texttt{creg} regression modeling framework to train a 11-class logistic regression classifier.\footnote{https://github.com/redpony/creg}  We parameterize the classifier as a multiclass logistic regression:
\begin{align*}
p_{\boldsymbol{\lambda}}(y& \mid \textbf{x})= \frac{\exp \sum_j \lambda_j h_j(\textbf{x},y)}{Z_{\boldsymbol{\lambda}}(\textbf{x})} ,
\end{align*}
where $\textbf{x}$ are documents, $h_j(\cdot)$ are real-valued feature functions of the document being classified, $\lambda_j$ are the corresponding weights, and $y$ is one of the eleven L1 class labels. To train the parameters of our model, we minimized the following objective,
\begin{align*}
\mathcal{L} = \alpha \overbrace{\sum_j \lambda_j^2}^{\textrm{$\ell_2$ reg.}} - \sum_{\{(\textbf{x}_i,y_i )\}_{i=1}^{|\mathcal{D}|}} \Big(& \overbrace{\log p_{\boldsymbol{\lambda}}(y_i \mid \textbf{x}_i)}^{\textrm{log likelihood}}  + \\
& \tau \underbrace{\mathbb{E}_{p_{\boldsymbol{\lambda}}(y' \mid \textbf{x}_i)} \log p_{\boldsymbol{\lambda}}(y' \mid \textbf{x}_i)}_{-\textrm{conditional entropy}}\Big) ,
\end{align*}
which combines the negative log likelihood of the training dataset $\mathcal{D}$, an $\ell_2$ (quadratic) penalty on the magnitude of $\boldsymbol{\lambda}$ (weighted by $\alpha$), and the \emph{negative} entropy of the predictive model (weighted by $\tau$). While an $\ell_2$ weight penalty is standard in regression problems like this, we found that the the additional entropy term gave more reliable results. Intuitively, the entropic regularizer encourages the remain maximally uncertain about its predictions. In the metaphor of ``maximum entropy'', the entropic prior finds a solution that has \emph{more} entropy than the ``maximum'' model that is compatible with the constraints.

@@

  \textnl{because}, \textnl{in fact}, etc.) Translators tend to spell out implicit

@@

  (\textnl{indicate}, \textnl{imply}, \textnl{contain}, etc.)

@@

\item[Misspelling features] Learning to perceive, produce, and encode non-native phonemic contrasts is extremely difficult for L2 learners \citep{hayes-harb:2008}. Since English's orthography is largely phonemic---even if it is irregular in many places, we expect leaners whose native phoneme contrasts are different from those of English to make characteristic spelling errors. For example, since Japanese and Korean lack a phonemic /l/-/r/ contrast, we expect native speakers of those languages to be more likely to make spelling errors that confuse {\tt l} and {\tt r} relative to native speakers of languages such as Spanish in which that pair is contrastive. To make this information available to our model, we use a noisy channel spelling corrector \citep{kernighan:1990} to identify and correct misspelled words in the training and test data. From these corrections, we extract minimal edit features that show what insertions, deletions, substitutions and joinings (where two separate words are written merged into a single orthographic token) were made by the author of the essay.

@@

\begin{compactdesc}
\item[\feat{DocLen}] Document length in tokens.
\item[\feat{Punct}] Counts of each punctuation mark. 
\item[\feat{Pron}] Counts of each pronoun.
\item[\feat{Position}] Positional token frequency. We use the counts
  for the first two and last three words before the period in each
  sentence as features. High-weight features for the \emph{second} word include: {\tt ARA:2<,>}; {\tt CHI:2<is>}; {\tt HIN:2<can>}.
\item[\feat{PsvRatio}] The proportion of passive verbs out of all
  verbs.
\item[\feat{CxtFxn}] Contextual function words. \raggedright High-weight features include: {\tt CHI:<some JJ>}; \mbox{\tt HIN:<as VBN>}. 
\item[\feat{Misspell}] Spelling correction edits. Features included substitutions, deletions, insertions, doubling of letters and missing doublings of letters, and splittings (\textnl{alot}$\rightarrow$\textnl{a lot}), as well as the word position where the error occurred. \raggedright High-weight features include: {\tt ARA:DEL<e>}, {\tt ARA:INS<e>}, {\tt ARA:SUBST<e>/<i>}; {\tt GER:SUBST<z>/<y>}; {\tt JPN:SUBST<l>/<r>}, {\tt JPN:SUBST<r>/<l>}; {\tt SPA:DOUBLE<s>}, {\tt SPA:MID\_INS<s>}, {\tt SPA:INS<s>}.
\item[\feat{Restore}] Counts of substitutions, deletions and
  insertions of predefined tokens that we restored in the texts. High-weight features include: {\tt CHI:DELWORD<do>}; {\tt GER:DELWORD<on>}; {\tt ITA:DELWORD<be>}  
\end{compactdesc}
\noindent
\Tref{tbl:addfeats} reports the empirical improvement that each of
these brings independently when added to the main features
(\Sref{sec:mainfeats}).

@@

Production of L2 texts, not unlike translating from L1 to L2, involves
a tension between the imposing models of L1 (and the source text), on
the one hand, and a set of cognitive constraints resulting from the
efforts to generate the target text, on the other. The former is
called \emph{interference} in Translation Studies \citep{Toury:1995}
and \emph{interlanguage} in second language acquisition
\citep{Selinker1972}. \citet{vered:noam:shuly} designed 32 classifiers to test the validity
of the forces acting on translated texts, and found that features sensitive to interference
consistently yielded the best performing classifiers. And indeed, in
this work too, we find fingerprints of the source language are dominant in the
makeup of L2 texts. The main difference, however, between texts translated by
professionals and the texts we address here, is that more often than
not professional translators translate into their mother tongue,
whereas L2 writers write out of their mother tongue by definition. So
interference is ever more exaggerated in this case, for example, also
phonologically \citep{tsur-rappoport:2007:Cognitive-2007}.

We explore the effects of interference by analyzing several patterns we observe in the features.
Our classifier finds that the
character sequence \textnl{alot} is overrepresented in Arabic L2
texts. Arabic has no indefinite article and we speculate that Arabic
speakers conceive \textnl{a lot} as a single word; the Arabic
equivalent for \textnl{a lot} is used adverbially like an \textnl{-ly}
suffix in English. For the same reason, another prominent feature is a
missing definite article before nouns and adjectives. Additionally,
Arabic, being an Abjad language, rarely indicates vowels, and indeed
we find many missing \textnl{e}-s and \textnl{i}-s in the texts of
Arabic speakers. Phonologically, because Arabic conflates
/\textipa{I}/ and /\textipa{@}/ into /i/ (at least in Modern Standard
Arabic), we see that many {\tt e}'s are indeed substituted for
{\tt i}'s in these texts.

@@

Other informative features that reflect L1 features include frequent misspellings involving confusions of {\tt l} and {\tt r} in Japanese essays. More mysteriously, the characters \textnl{r} and \textnl{s} are misused in Chinese
and Spanish, respectively. The word \textnl{then} is dominant in the texts of
Hindi speakers. Finally, it is clear that authors refer to their
native cultures (and, consequently, native languages and countries);
the strings \textnl{Turkish}, \textnl{Korea}, and \textnl{Ita} were
dominant in the texts of Turkish, Korean and Italian native speakers,
respectively.

@@

Our point of departure was the analogy between translated texts as a genre in its own and L2 writers as pseudo translators, relying heavily on their mother tongue and transferring their native models to a second language. In formulating our features, we assumed that like translators, L2 writers will write in a simplified manner and overuse explicit markers. Although this should be studied vis-a-vis comparable outputs of mother tongue writers in English, we observe that the best features of our classifiers are of the ``interference'' type, i.e. phonological, morphological and syntactic in nature, mostly in the form of misspelling features, restoration tags, punctuation and lexical and syntactic modeling.

@@

@@

\bibliography{l1id}
\bibliographystyle{plainnat}

>>>>

%\usepackage[round]{natbib}
\newcommand{\citep}[1]{\cite{#1}}
\newcommand{\citet}[1]{\newcite{#1}}

\setlength\titlebox{6.5cm}    % Expanding the titlebox

@@

\newcommand{\feat}[1]{\textsmaller[.5]{\textsf{#1}}} % code for a feature group
\newcommand{\featstr}[1]{\textsmaller[.5]{\verb|#1|}} % full feature name
\newcommand{\textnl}{\textit}

@@

\section{Related work}
\label{sec:literature}
The task of L1 identification was introduced by Koppel et~al. \shortcite{koppel2005automatically,koppel2005determining}%
%\citet{koppel2005automatically,koppel2005determining}
, who work on the
International Corpus of Learner English \citep{icle}, which includes
texts written by students from 5 countries, Russia, the Czech Republic, Bulgaria,
France, and Spain. The texts range from 500~to 850~words in length. Their
classification method is a linear SVM, and features include 400~standard function words, 200~letter $n$-grams, 185~error types and 250
rare part-of-speech (POS) bigrams. Ten-fold cross-validation results
on this dataset are 80\% accuracy.

The same experimental setup is assumed by
\citet{tsur-rappoport:2007:Cognitive-2007}, who are mostly interested
in testing the hypothesis that an author's choice of words in a second
language is influenced by the \emph{phonology} of his or her L1. They confirm
this hypothesis by carefully analyzing the features used by
Koppel et~al.\nocite{koppel2005automatically}%\citeauthor{koppel2005automatically}
, controlling for potential
biases.

Wong and Dras \shortcite{Wong-Dras:2009:ALTA2009,wong-dras:2011:EMNLP}
%\citet{Wong-Dras:2009:ALTA2009,wong-dras:2011:EMNLP} 
are also
motivated by a linguistic hypothesis, namely that \emph{syntactic} errors in
a text are influenced by the author's
L1. \citet{Wong-Dras:2009:ALTA2009} analyze three error types
statistically, and then add them as features in the same experimental
setup as above (using LIBSVM with a radial kernel for
classification). The error types are subject-verb disagreement,
noun-number disagreement and misuse of determiners. Addition of these
features does not improve on the results of
Koppel et~al.\nocite{koppel2005automatically}%\citeauthor{koppel2005automatically}
. \citet{wong-dras:2011:EMNLP}
further extend this work by adding as features horizontal slices of
parse trees, thereby capturing more syntactic structure. This improves
the results significantly, yielding 78\% accuracy compared with less
than 65\% using only lexical features.

@@

\section{Model}
\label{sec:model}
For our classification model we used the \texttt{creg} regression modeling framework to train a 11-class logistic regression classifier.\footnote{\url{https://github.com/redpony/creg}}  We parameterize the classifier as a multiclass logistic regression:
\begin{align*}
p_{\boldsymbol{\lambda}}(y& \mid \textbf{x})= \frac{\exp \sum_j \lambda_j h_j(\textbf{x},y)}{Z_{\boldsymbol{\lambda}}(\textbf{x})} ,
\end{align*}
where $\textbf{x}$ are documents, $h_j(\cdot)$ are real-valued feature functions of the document being classified, $\lambda_j$ are the corresponding weights, and $y$ is one of the eleven L1 class labels. To train the parameters of our model, we minimized the following objective,
\begin{align*}
\mathcal{L} = \alpha \overbrace{\sum_j \lambda_j^2}^{\textrm{$\ell_2$ reg.}} - \sum_{\{(\textbf{x}_i,y_i )\}_{i=1}^{|\mathcal{D}|}} \Big(& \overbrace{\log p_{\boldsymbol{\lambda}}(y_i \mid \textbf{x}_i)}^{\textrm{log likelihood}}  + \\
& \tau \underbrace{\mathbb{E}_{p_{\boldsymbol{\lambda}}(y' \mid \textbf{x}_i)} \log p_{\boldsymbol{\lambda}}(y' \mid \textbf{x}_i)}_{-\textrm{conditional entropy}}\Big) ,
\end{align*}
which combines the negative log likelihood of the training dataset $\mathcal{D}$, an $\ell_2$ (quadratic) penalty on the magnitude of $\boldsymbol{\lambda}$ (weighted by $\alpha$), and the \emph{negative} entropy of the predictive model (weighted by $\tau$). While an $\ell_2$ weight penalty is standard in regression problems like this, we found that the the additional entropy term gave more reliable results. Intuitively, the entropic regularizer encourages the model to remain maximally uncertain about its predictions. In the metaphor of ``maximum entropy'', the entropic prior finds a solution that has \emph{more} entropy than the ``maximum'' model that is compatible with the constraints.

@@

  \textnl{because}, \textnl{in fact}, etc.). Translators tend to spell out implicit

@@

  (\textnl{indicate}, \textnl{imply}, \textnl{contain}, etc.).

@@

\item[Misspelling features] Learning to perceive, produce, and encode non-native phonemic contrasts is extremely difficult for L2 learners \citep{hayes-harb:2008}. Since English's orthography is largely phonemic---even if it is irregular in many places, we expect leaners whose native phoneme contrasts are different from those of English to make characteristic spelling errors. For example, since Japanese and Korean lack a phonemic /l/-/r/ contrast, we expect native speakers of those languages to be more likely to make spelling errors that confuse \textnl{l} and \textnl{r} relative to native speakers of languages such as Spanish in which that pair is contrastive. To make this information available to our model, we use a noisy channel spelling corrector \citep{kernighan:1990} to identify and correct misspelled words in the training and test data. From these corrections, we extract minimal edit features that show what insertions, deletions, substitutions and joinings (where two separate words are written merged into a single orthographic token) were made by the author of the essay.

@@

\begin{compactdesc}
\item[\feat{DocLen}] Document length in tokens.
\item[\feat{Punct}] Counts of each punctuation mark. 
\item[\feat{Pron}] Counts of each pronoun.
\item[\feat{Position}] Positional token frequency. We use the counts
  for the first two and last three words before the period in each
  sentence as features. High-weight features for the \emph{second} word include: {\tt ARA:2<,>}; {\tt CHI:2<is>}; {\tt HIN:2<can>}.
\item[\feat{PsvRatio}] The proportion of passive verbs out of all
  verbs.
\item[\feat{CxtFxn}] Contextual function words. \raggedright High-weight features include: {\tt CHI:<some JJ>}; \mbox{\tt HIN:<as VBN>}. 
\item[\feat{Misspell}] Spelling correction edits. Features included substitutions, deletions, insertions, doubling of letters and missing doublings of letters, and splittings (\textnl{alot}$\rightarrow$\textnl{a lot}), as well as the word position where the error occurred\nss{i.e. beginning, middle, or end?}. \raggedright High-weight features include: {\tt ARA:DEL<e>}, {\tt ARA:INS<e>}, {\tt ARA:SUBST<e>/<i>}; {\tt GER:SUBST<z>/<y>}; {\tt JPN:SUBST<l>/<r>}, {\tt JPN:SUBST<r>/<l>}; {\tt SPA:DOUBLE<s>}, {\tt SPA:MID\_INS<s>}, {\tt SPA:INS<s>}.
\item[\feat{Restore}] Counts of substitutions, deletions and
  insertions of predefined tokens that we restored in the texts. High-weight features include: {\tt CHI:DELWORD<do>}; {\tt GER:DELWORD<on>}; {\tt ITA:DELWORD<be>}  
\end{compactdesc}
\noindent
\Tref{tbl:addfeats} reports the empirical improvement that each of
these brings independently when added to the main features
(\Sref{sec:mainfeats}).

@@

Production of L2 texts, not unlike translating from L1 to L2, involves
a tension between the imposing models of L1 (and the source text), on
the one hand, and a set of cognitive constraints resulting from the
efforts to generate the target text, on the other. The former is
called \emph{interference} in Translation Studies \citep{Toury:1995}
and \emph{interlanguage}\nss{do you mean `transfer'? I think `interlanguage' refers to the L2 as used by a nonnative speaker, and `transfer' from the L1 is one of the influences on interlanguage.} in second language acquisition
\citep{Selinker1972}. \citet{vered:noam:shuly} designed 32 classifiers to test the validity
of the forces acting on translated texts, and found that features sensitive to interference
consistently yielded the best performing classifiers. And indeed, in
this work too, we find fingerprints of the source language are dominant in the
makeup of L2 texts. The main difference, however, between texts translated by
professionals and the texts we address here, is that more often than
not professional translators translate into their mother tongue,
whereas L2 writers write out of their mother tongue by definition. So
interference is ever more exaggerated in this case, for example, also
phonologically \citep{tsur-rappoport:2007:Cognitive-2007}.

We explore the effects of interference by analyzing several patterns we observe in the features.
Our classifier finds that the
character sequence \textnl{alot} is overrepresented in Arabic L2
texts. Arabic has no indefinite article and we speculate that Arabic
speakers conceive \textnl{a lot} as a single word; the Arabic
equivalent for \textnl{a lot} is used adverbially like an \textnl{-ly}
suffix in English. For the same reason, another prominent feature is a
missing definite article before nouns and adjectives. Additionally,
Arabic, being an Abjad language, rarely indicates vowels, and indeed
we find many missing \textnl{e}'s and \textnl{i}'s in the texts of
Arabic speakers. Phonologically, because Arabic conflates
/\textipa{I}/ and /\textipa{@}/ into /i/ (at least in Modern Standard
Arabic), we see that many \textnl{e}'s are indeed substituted for
\textnl{i}'s in these texts.

@@

Other informative features that reflect L1 features include frequent misspellings involving confusions of \textnl{l} and \textnl{r} in Japanese essays. More mysteriously, the characters \textnl{r} and \textnl{s} are misused in Chinese
and Spanish, respectively. The word \textnl{then} is dominant in the texts of
Hindi speakers. Finally, it is clear that authors refer to their
native cultures (and, consequently, native languages and countries);
the strings \textnl{Turkish}, \textnl{Korea}, and \textnl{Ita} were
dominant in the texts of Turkish, Korean and Italian native speakers,
respectively.

@@

Our point of departure was the analogy between translated texts as a genre in its own and L2 writers as pseudo translators, relying heavily on their mother tongue and transferring their native models to a second language. In formulating our features, we assumed that like translators, L2 writers will write in a simplified manner and overuse explicit markers. Although this should be studied vis-\`{a}-vis comparable outputs of mother tongue writers in English, we observe that the best features of our classifiers are of the ``interference'' type, i.e. phonological, morphological and syntactic in nature, mostly in the form of misspelling features, restoration tags, punctuation and lexical and syntactic modeling.

@@

\nss{comment on how we did in relation to other task participants? if I were a reader I'd want to know}

@@

\bibliography{l1id}
%\bibliographystyle{plainnat}
\bibliographystyle{naaclhlt2013}
